---
title: "IS607 Project 1"
author: "FD"
output: html_document
---

Load the datafile, skipping the first three junk lines so we can begin at the first row of data.


```{r, warning=FALSE, message=FALSE}
chess <- read.csv("https://raw.githubusercontent.com/fdsps/IS607/master/Project1/tournamentinfo.txt", skip=3)
require(stringr)
```

Player data spans two rows, separated by a dashed line. We could strip out junk lines, or just use indices to focus on the rows we want, ignoring the junk.


```{r}
toprow<-seq(1,nrow(chess),3)
botrow<-seq(2,nrow(chess),3)
```

From here, we just create new lists of data using regex.

```{r}
opps <-str_extract_all(chess[toprow,1],"\\d{1,2}(?=\\|)") # collect the opposing player IDs for use later.

players<- str_extract_all(chess[toprow,],"(?<=\\| )(\\w|\\s)+(?=\\|)")

totpts <-str_extract_all(chess[toprow,1],"\\d\\.\\d")

prerat <- as.numeric(str_extract_all(chess[botrow,1],"(?<=R:)\\s+\\d+")) # as.numeric for use in calculating the mean

state <- str_sub(chess[botrow,1], start = 4, end = 5) # always in same position.
```

Now to calculate the opposing player averages. We use each row of *opps* as a list of keys into *prerat* (prerating), using sapply to calculate the mean on each row, conveniently saved to a list.

```{r}
opps_avg <- sapply(opps, function(x) mean(prerat[as.numeric(x)]))
```

Finally, build the dataframe with some clean-up.

```{r}
df<- data.frame("players" = unlist(str_trim(players)), "state" = state, "totpts" = unlist(totpts), "prerat" = unlist(prerat), 
"opps_avg"= round(opps_avg) )
df
```

