---
title: "IS607_wk4_assmnt"
author: "FDSPS"
output: html_document
---

p. 217

a) [0-9]+\\$

Match a string with one or more digits followed by $.
```{r}
grep("[0-9]+\\$","100$") 
grep("[0-9]+\\$","foo$") # fails
grep("[0-9]+\\$","100") # fails
grep("[0-9]+\\$","foo2$")
```

b) \\b[a-z]{1,4}\\b

Match words 1-4 chars in length of alphabetic characters.

```{r}
grep("\\b[a-z]{1,4}\\b","a b")
grep("\\b[a-z]{1,4}\\b","5 abcd 5") 
grep("\\b[a-z]{1,4}\\b","5 abcde 5") # fails
```

c) .*?\\\\.txt$

Match a string ending in .txt (eg. find text files).

```{r}
grep(".*?\\.txt$","hello world.txt")
grep(".*?\\.txt$","hello world.dat") # fail
grep(".*?\\.txt$","hello world.dat.txt")
```

The ? is redundant, I think.

d) \\\\d{2}/\\\\d{2}/\\\\d{4}

Date string in the format dd/dd/dddd

```{r}
grep("\\d{2}/\\d{2}/\\d{4}","11/22/2015")
grep("\\d{2}/\\d{2}/\\d{4}","11/22/15") # nah
```


e) <(.+?)>.+?<\/\\\\1>

Match balanced XML <.>.<\/.> tags with no whitespace in or between using backreference.

```{r}
grep("<(.+?)>.+?</\\1>","<f>foo</f>")
grep("<(.+?)>.+?</\\1>","<f>foo</g>") # no
```

5.

[^a-zA-Z]{1,}\\\\$

6.

a)
```{r}
library(stringr)
s <- "chunkylover53[at]aol[dot]com"
str_replace(s, "(.+)\\[at\\](.+)\\[dot\\](com)", "\\1@\\2.\\3")
```

b) [:digit:] matches only the first digit. Use [:digit:]+

c) \\\\D matches first non-digit. Use \\\\d+

```{r}
str_extract(s, "[:digit:]")
str_extract(s, "[:digit:]+")
str_extract(s, "\\D")
str_extract(s, "\\d+")
```
